Learning Actionable Analytics (with applications for reducing defects and reducing runtimes)

### 2015-05-14 how.tex
Updates from ShareLaTeX

 This paper proposes one such planner. ``HOW'' is a 
 tool for proposing changes to independent
 variables in order to improve on 
 the predictions of the dependent variables. Unlike other approaches
 for learning optimizations to software, HOW does not require
 an underlying of the domain. Also, HOW's plans
 are designed to result in {\em minimal displacement}
 to current practice.

@@

\section{Introduction}
Suppose someone suggests a change to your project. How might you assess the effectiveness of that change?

This is an open and pressing issue.
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself; e.g.
\bi
\item 
Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
How to assess such {\em plans}?
Some organizations have the resources to 
run repeated trials to assess  different treatments.
For example, in one remarkable recent study, Bente et al. report results
were the same specification was developed  by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
to (say) take a code base, assign it to $N$ different teams, apply
HOW's plans the code used by  $N/2$ teams, then check in 12 months time
if those halves of the teams have fewer defects than the other.

Very few industrial or research groups have access
to the kinds of resources using by Bente et al. (evidence: in the size years since the
publication of that work, we know of only one of two similar studies). Also, given the
diversity of modern software projects, it might be unreasonable to demand that all
proposed changes for all projects are always evaluated by something like the Bente et al. study.

This paper proposes another, less resource intensive,  method for assessing the effectiveness of proposed
changes to a project.  
We propose using prediction systems (built by data mining) as a verification
tool for planners. In domains where  data miners can build
predictors of adequate performance, then those predictors can assess the value
of proposed changes.

The planner explored in this paper is the ``HOW'' case-based reasoner (CBR).
HOW's  ``plans'' are  set of changes
proposed  in order to better achieve some goal. 
Specifically, for every test case $Z$, HOW proposes a  plan $\delta$ to
  adjust   attributes $Z_j$:
\begin{equation}\label{eq:one}
Z_j = Z_j + \delta_j
\end{equation}
For example, to simplify a  large bug-prone  method, HOW might suggest
to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
two simpler functions).

HOW's  plans displace the test case into some region
of the training data. This is important since we will use that training data
to build a predictor to assess the impact of that plan.
To put that another way, when a novel test case is presented,
HOW says 
\begin{quote}{\em ``I may not know too much about that, but once you change into something like
what I've already seen, then I can suggest how to change it''.}
\end{quote}
This paper shows experimentally  that   HOW's plans can be quite effective and 
 significantly reduce the  predicted
value of:
\bi
\item The  defects found  in Jureczko et al.'s    object-oriented JAVA systems~\cite{jureczko10}.
\item The runtime of the  software systems   configured by  Seigmund et al.~\cite{sven12}.
\ei 
The rest of this paper presents the background to this work in Section~2; the design of HOW in Section~3;
then our experimental results  in Section~4. Those experiments show that HOW can   reduce the expected
value of defects in Jureczko   data sets by 72\% to 66\%  (median,best).
Also, in the Seigmund data sets, HOW can   reduce the expected value of the runtimes by  75\% to 20\% (median,best).

@@

\begin{figure}[!t] 
  {\scriptsize \textbf{Ant}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~~~~}c@{}r}}
      \arrayrulecolor{darkgray}
      \rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} && \textbf{Bugs}\\
   1 &  F=0.75, B=100 &    0.71  &  0.19 & \quart{0}{25}{10}{49} & 62 \\
\hline  2 &   F=0.5, B=100  &    0.88  &  0.14 & \quart{20}{19}{33}{49} & 77\\
\hline  3 & F=0.5, B=50\% &    0.91  &  0.08 & \quart{32}{11}{37}{49} & 80\\
  3 & F=0.75, B=50\% &    0.92  &  0.09 & \quart{31}{12}{39}{49} & 82\\
  3 &  F=0.25, B=100  &    0.92  &  0.04 & \quart{36}{5}{39}{49} & 81\\
  3 & F=0.75, B=25\% &    0.94  &  0.09 & \quart{36}{12}{41}{49} & 82\\
\hline  4 & F=0.25, B=50\% &    0.97  &  0.03 & \quart{44}{4}{45}{49} & 86\\
  4 & F=0.5, B=25\% &    0.98  &  0.06 & \quart{40}{8}{47}{49} & 86\\
\hline  5 &F= 0.25, B=25\% &    0.99  &  0.01 & \quart{48}{1}{48}{49} & 87\\
\hline  6 &         Base &    1.0  &  0.0 & \quart{49}{0}{49}{49} & 90\\
\hline \end{tabular}}
%\end{minipage}\\[0.1cm]

@@

%XXX begin rahul
One way to assess the validity of our predictions is to consider the space that generates
them. \fig{howxy} shows (1)~the training cases and (2)~the test cases and (3)~the
test cases displaced after applying HOW's plans for the {\em ant} data set.
In this figure, all the cases are projected onto two dimensions by   finding their $x,y$ positions in the top-level {\tt cluster},
as generated by \fig{howcode}. Note that the  the   displaced test
cases   fall into nearly the same regions as the training data.

This pattern of \fig{howxy} (where the displaced test cases are found in the same
region as the training data) has been observed in all the other data sets studied in this
paper. Hence, the
predictors learned from thse training data have some authority in the regions
used to generate our predictions regarding the effects of planning\footnote{If the reader is concerned that this 2d projection
losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
have confirmed that building defect predictors from this 2d space are just as effective
as predictors learned using all the defect attributes of \fig{ck}.}.

\begin{figure}
\begin{center}\includegraphics[width=3in]{figs/2d.png}\end{center}
\caption{Training cases (in gray), test cases (in red) and 
test cases displaced by HOW (in green). Note how the green and the gray occupy
the same regions; i.e. predictors trained on the training cases can be applied
to the displaced test cases.}\label{fig:howxy}
\end{figure}
%XXX end

>>>>

 This paper proposes one such planner. ``HOW'' is a 
 tool for proposing changes to independent
 variables in order to improve on 
 the predictions of the dependent variables. Unlike other approaches
 for learning optimizations to software, HOW does not require
 an underlying model of the domain. Also, HOW's plans
 are designed to result in {\em minimal displacement}
 to current practice.

@@

\section{Introduction}
Suppose someone wants changes that can improve their software project. How would you propose
sensible changes? And how would you assess the effectiveness of those changes?

This is an open and pressing issue.
Business  users   now demand   data mining tools
that  support a  business-level
interpretation of that data. For example,
at a  panel on software analytics at ICSE'12,
industrial practitioners lamented the state of the art in data mining
and software engineering~\cite{menzies12a}. Panelists commented that
``prediction is all well and good, but what about decision
making?''. That is, these panelists are more interested in the interpretations
and follow-up
that occurs {\em after} the mining, rather than just  the mining itself; e.g.
\bi
\item 
Instead of just accepting  {\em predictions} on how many 
 software defects
to expect,  business users might now demand a {\em plan} to
reduce the likelihood of those defects;
\item Instead of just accepting {\em predictions} on the runtime
time of their software, business users might now demand
a {\em plan} to reduce that runtime.
\ei
To generate those plans, we propose the ``HOW'' case-based reasoner (CBR).
HOW's  ``plans'' are  set of changes
proposed  in order to better achieve some goal. 
Specifically, for every test case $Z$, HOW proposes a  plan $\delta$ to
  adjust   attributes $Z_j$:
\[
Z_j = Z_j + \delta_j
\]
For example, to simplify a  large bug-prone  method, HOW might suggest
to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
two simpler functions).
HOW's  plans displace the test case into some region
of the training data. This is important since we will use that training data
to build a predictor to assess the impact of that plan.
To put that another way, when a novel test case is presented,
HOW says 
\begin{quote}{\em ``I may not know too much about that, but once you change into something like
what I've already seen, then I can suggest how to improve it''.}
\end{quote}
After generating the plans, the next question is ``how to assess them?''.  
Some organizations have the resources to 
run repeated trials to assess  different treatments.
For example, in one remarkable recent study, Bente et al. report results
were the same specification was developed  by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
to (say) take a code base, assign it to different teams, make these teams  adopt different polices,
then check in 12 months time
 which teams have fewer defects than the others.

Very few industrial or research groups have access
to the kinds of resources needed for this kind study  (evidence: in the six years since the
publication of that work, we know of only one   similar study to Bente et al.). Also, given the
diversity of modern software projects, it might be unreasonable to demand that all
proposed changes for all projects are always evaluated by something like the Bente et al. study.

This paper proposes another, less resource intensive,  method for assessing the effectiveness of proposed
changes to a project.  
We  use a prediction system (built by data mining) as a verification
tool for planners. In domains where  data miners can build
predictors of adequate performance, then those predictors can assess the value
of proposed changes.
This paper applies this proposal to show that   HOW's plans are quite effective and  can
 significantly reduce the  predicted
value of:
\bi
\item The  defects found  in Jureczko et al.'s    object-oriented JAVA systems~\cite{jureczko10}.
\item The runtime of the  software systems   configured by  Seigmund et al.~\cite{sven12}.
\ei 
The rest of this paper presents the background to this work in Section~2; the design of HOW in Section~3;
then our experimental results  in Section~4.  Our conclusion is that we can plan for the future in
software engineering, provided the domain data supports adequate predictors.

@@

\begin{figure}[!t] 
  {\scriptsize \textbf{Ant}\\[0.1cm]}
  {\scriptsize  \begin{tabular}{{l@{~~~~}l@{}r@{~~~~}r@{~~}c@{}r}}
      \arrayrulecolor{darkgray}
      \rowcolor[gray]{.9} \textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} && \textbf{Bugs}\\
   1 &  F=0.75, B=100 &    0.71  &  0.19 & \quart{0}{25}{10}{49} & 62 \\
\hline  2 &   F=0.5, B=100  &    0.88  &  0.14 & \quart{20}{19}{33}{49} & 77\\
\hline  3 & F=0.5, B=50\% &    0.91  &  0.08 & \quart{32}{11}{37}{49} & 80\\
  3 & F=0.75, B=50\% &    0.92  &  0.09 & \quart{31}{12}{39}{49} & 82\\
  3 &  F=0.25, B=100  &    0.92  &  0.04 & \quart{36}{5}{39}{49} & 81\\
  3 & F=0.75, B=25\% &    0.94  &  0.09 & \quart{36}{12}{41}{49} & 82\\
\hline  4 & F=0.25, B=50\% &    0.97  &  0.03 & \quart{44}{4}{45}{49} & 86\\
  4 & F=0.5, B=25\% &    0.98  &  0.06 & \quart{40}{8}{47}{49} & 86\\
\hline  5 &F= 0.25, B=25\% &    0.99  &  0.01 & \quart{48}{1}{48}{49} & 87\\
\hline  6 &         Base &    1.0  &  0.0 & \quart{49}{0}{49}{49} & 90\\
\hline \end{tabular}}
%\end{minipage}\\[0.1cm]

@@

One way to assess the validity of our predictions is to consider the space that generates
them. \fig{howxy} is from the {\em ivy} data
set. It shows (1)~the training cases and (2)~the test cases and (3)~the
test cases displaced after applying HOW's plans.
To generate that figure, we placed all of 1+2+3 into the same data set then computed the first
two  components (using PCA). Note that the  the   displaced test
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).  That is, if we believe
that the gray data is  suitable for assessing
the test cases (as indicated by the {\em ivy} results from \fig{j}), then it follows
that we believe even more in the predictions
generated from the displaced data).

This pattern of \fig{howxy} (where the displaced test cases are found in the same
region as the training data) has been observed in all the other data sets studied in this
paper. Hence, the
predictors learned from these training data have some authority in the regions
used to generate our predictions regarding the effects of planning\footnote{If the reader is concerned that this 2d projection
losses important information, then they should note that two separate  studies~\cite{papa13,divya15}
have confirmed that building defect predictors from this 2d space are just as effective
as predictors learned using all the defect attributes of \fig{ck}.}.

\begin{figure}
\begin{center}\includegraphics[width=3in]{figs/2d.png}\end{center}
\caption{Training cases (in gray), test cases (in red) and 
test cases displaced by HOW (in green). Note how the green (displaced) cases are closer to the  training
date than the original test data
i.e. predictors trained on the training cases and
applied to the test cases should
work just as well, or better, on the displaced test cases.}\label{fig:howxy}
\end{figure}
%XXX end
